Global patterns of forest autotrophic carbon fluxes

### 2020-01-17 manuscript/manuscript draft.Rmd
working on Helene's comments

**Forests play a critical role in regulating atmospheric CO~2~ and climate** (**Bonan**). Approximately **##** Gt of CO~2~ cycle through Earth's forests each year (*i.e.*, total gross primary production, $GPP$; **REF-KAT**), and their net C sequestration (~2.4 GtCyr^-1^) offsets roughly 30% of anthropogenic fossil fuel emissions [@pan_large_2011]. As atmospheric carbon dioxide levels continue to rise, driving the global climate crisis, there is increasing recognition that  protection and management of forest resources will play an important role in mitigating climate change (** REFS- e.g., IPCC 1.5C and land reports**). Carbon fixation, allocation, and metabolism by trees set the basis for energy and material flows in forest ecosystems and define their interactions with Earth's changing climate.

@@

**The recent development of a global forest carbon database synthesizing multiple variables and including records of stand history (ForC; [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc:_2018]) opens up the possibility for a standardized analysis of global scale variation** in multiple components of forest productivity and the principle climatic drivers of these patterns.  In order to approach these broad and complex issues, we simplify the major gaps in our knowledge to five broad hypotheses and corresponding specific predictions (Table 1). First, we ask how forest autotrophic carbon fluxes (FACF) vary with latitude. We then test how these fluxes relate to MAT and MAP, and additionally how they respond to other, less well studied, climate variables. We finally consider the relationship between FACF and seasonality, considering the role of seasonality in explaining variation in carbon fluxes, and the influence of climate on FACF per month of the growing season. We use a comprehensive global database of forest carbon fluxes to explore the above questions for nine carbon fluxes, allowing for an in-depth exploration of the effect of climate on global productivity.

@@

| Variable | Definition | Components included | Methodologies used | Number of records | Number of geographic areas |
|--------------|-------------------|-------------|------------------------|-------|-------|
| $GPP$ | Annual gross primary production; annual uptake of carbon dioxide by an ecosystem | NA | Flux partitioning of eddy covariance | 243 | 49 |
| $NPP$ | Annual net primary production; the component of GPP that is stored in plant tissue; $GPP - R_{auto}$ | Foliage, branch, stem, coarse root and fine root | Direct measurement of annual increments of components | 92 | 42 |
| $ANPP$ | Aboveground net primary production | Foliage, stem, and optionally branch | Direct measurement of annual increments of components | 256 | 80 |
| $ANPP_{foliage}$ | Net primary production of foliage | Foliage | Direct measurement of litterfall, correcting for changes in leaf biomass when measured | 98 | 49 |
| $ANPP_{woody-stem}$ | Net primary production of woody stems | Woody stems | Direct measurement of stem growth increment | 264 | 96 |
| $BNPP_{root}$  | Belowground net primary production | Coarse and fine roots |  Coarse roots estimated indirectly using allometries based on aboveground stem increment measures ; fine roots as below | 101 | 48 |
| $BNPP_{fine.root}$ | Net primary production of fine roots | Fine roots | Measurements combined one or more of the following: minirhizotrons, fine root turnover, soil cores, root ingrowth cores | 88 | 41 |
| $R_{auto}$ | Annual autotrophic respiration, including above- and belowground components | Foliage, stem, and root | Chamber measurements of component gas exchange | 22 | 13 |
| $R_{root}$ | Annual root respiration | Coarse and fine roots | *Measurement of root gas exchange, root exclusion from soil respiration chambers* | 64 | 26 |

>>>>

**Forests play a critical role in regulating atmospheric CO~2~ and climate** (**Bonan**). Approximately **##** Gt of CO~2~ cycle through Earth's forests each year **REF-KAT**), and their net C sequestration (~2.4 GtCyr^-1^) offsets roughly 30% of anthropogenic fossil fuel emissions [@pan_large_2011]. As atmospheric carbon dioxide levels continue to rise, driving climate change, there is increasing recognition that  protection and management of forest resources will play an important role in mitigating climate change (** REFS- e.g., IPCC 1.5C and land reports**). Carbon fixation, allocation, and metabolism by trees set the basis for energy and material flows in forest ecosystems and define their interactions with Earth's changing climate.

@@

**The recent development of a global forest carbon database synthesizing multiple variables and including records of stand history (ForC; [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc:_2018]) opens up the possibility for a standardized analysis of global scale variation** in multiple components of forest productivity and the principle climatic drivers of these patterns.  In order to approach these broad and complex issues, we simplify the major gaps in our knowledge to five broad hypotheses and corresponding specific predictions (Table 1). ** see [issue #48 ](https://github.com/forc-db/Global_Productivity/issues/48)** First, we ask how forest autotrophic carbon fluxes (FACF) vary with latitude. We then test how these fluxes relate to MAT and MAP, and additionally how they respond to other, less well studied, climate variables. Finally, we consider the relationship between FACF and seasonality, considering the role of seasonality in explaining variation in carbon fluxes, and the influence of climate on FACF standardized by growing season length. We use a comprehensive global database of forest carbon fluxes to address the above questions for nine carbon fluxes, allowing for an in-depth exploration of the effect of climate on global productivity.

@@

| Variable | Definition | Components included | Methodologies used | Number of records | Number of geographic areas |
|--------------|-------------------|-------------|------------------------|-------|-------|
| $GPP$ | Annual gross primary production; annual uptake of carbon dioxide by an ecosystem | NA | Flux partitioning of eddy covariance | 243 | 49 |
| $NPP$ | Annual net primary production; the component of GPP that is stored in plant tissue; $GPP - R_{auto}$ | Foliage, branch, stem, coarse root and fine root | Direct measurement of annual increments of components | 92 | 42 |
| $ANPP$ | Aboveground net primary production | Foliage, stem, and optionally branch | Direct measurement of annual increments of components | 256 | 80 |
| $ANPP_{foliage}$ | Net primary production of foliage | Foliage | Direct measurement of litterfall, correcting for changes in leaf biomass when measured | 98 | 49 |
| $ANPP_{woody-stem}$ | Net primary production of woody stems | Woody stems | Direct measurement of stem growth increment (**with recruitment?**) | 264 | 96 |
| $BNPP_{root}$  | Belowground net primary production | Coarse and fine roots |  Coarse roots estimated indirectly using allometries based on aboveground stem increment measures ; fine roots as below | 101 | 48 |
| $BNPP_{fine.root}$ | Net primary production of fine roots | Fine roots | Measurements combined one or more of the following: minirhizotrons, fine root turnover, soil cores, root ingrowth cores | 88 | 41 |
| $R_{auto}$ | Annual autotrophic respiration, including above- and belowground components | Foliage, stem, and root | Chamber measurements of component gas exchange | 22 | 13 |
| $R_{root}$ | Annual root respiration | Coarse and fine roots | *Measurement of root gas exchange, root exclusion from soil respiration chambers* | 64 | 26 |
